As a cis-gendered, bisexual man and an international student in computer science, my personal and academic experiences have shaped my commitment to fostering inclusivity and belonging in academia.
% 
Growing up in a multicultural environment, I developed a deep appreciation for the diversity of language and the ways it connects people across boundaries.
% 
This cultural perspective informs my teaching and research, where I strive to create spaces that celebrate diverse voices and empower all students to succeed.

My teaching philosophy emphasizes equity and collaboration.
% 
As a Teaching Assistant, I managed a team of seven TAs to design homework assignments that were accessible and inclusive.
% 
For instance, in an assignment on Seagull LM, a humor-generation model, I introduced an innovative evaluation metric---the log-odds ratio between student outputs and a reference model.
% 
This approach encouraged students to focus on understanding principles rather than competing for performance, fostering a learning environment that prioritized exploration over pressure.
% 
Additionally, by providing timely and thorough responses on EdStem (answering over $700$ questions across various courses), I ensured that all students, regardless of their background or prior experience, had access to the support and guidance they needed to succeed.

These experiences reflect my broader understanding of the barriers that students from underrepresented groups often face in STEM.
% 
As an international student, I experienced the challenges of navigating new academic systems, adapting to cultural differences, and overcoming feelings of isolation.
% 
These struggles motivated me to mentor students facing similar barriers, sharing strategies for success and advocating for practices that ensure equitable opportunities.

Beyond the classroom, my research collaborations have reinforced my belief in the value of diversity.
% 
Working on projects like MambaByte, a byte-level language model, I witnessed how teams with varied perspectives drive creative problem-solving and innovation.
% 
This project, which explored challenges in scaling language models to handle long contexts efficiently, benefited from insights contributed by collaborators from different cultural and academic backgrounds.
% 
These experiences have strengthened my resolve to foster diverse teams that reflect the richness of the world we aim to model through our research.

At \thecollegeabbr, I aim to contribute to initiatives that advance diversity, equity, and inclusion by mentoring students, creating accessible learning resources, and promoting collaborative research.
% 
My personal background and academic journey have shown me the transformative power of inclusive spaces where individuals feel empowered to bring their unique perspectives.
% 
By integrating my passion for teaching and research, I hope to help shape a community that celebrates diversity and drives meaningful change.
